Supporting Information

### 2017-04-06 code/rmarkdown/supporting-information.Rmd
massive overhaul - packrat to manage r packages - reformat code according to lintr - fix book references

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
```

@@

```{r, include=FALSE}
## initialization
# load .rda file
checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)

## prelimary processing
# redefine file paths for rasters
for (i in seq_len(nlayers(cs1.bioclim.RAST)))
	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'raptr-manuscript')[[1]][[2]])

@@

The reliable formulation (RRAP) is a multi-objective optimisation problem.

@@

We generated simulated conservation planning data sets using varying number of features `r paste(benchmark.params.LST$number.of.features, collapse=',')` and planning units (paste(benchmark.params.LST$number.of.planning.units, collapse=',')). For a given number of planning units and features, the data was simulated using the following steps. First, the planning units were simulated as squares that were arranged in a grid. All planning unit costs were set to one. The spatial distribution of features were generated using binary, spatially autocorrelated random fields [using the \texttt{RandomFields R} package; @rXXX]. If features were associated with less than two occupied planning units, these features were set to occupy two randomly selected planning units. All occupied planning units were set to have a probability of occupancy of `r benchmark.params.LST[[MODE]]$occupancy.probability`. Similar to the simulated data sets used in the main text, a geographic attribute space was used for each feature. The features' demand points were set as the centroids of the planning units that they occupied, and were weighted by the units' probability of occupancy. All problems used amount-based targets of `r benchmark.params.LST[[MODE]]$amount.target*100` % and space-based targets of `r benchmark.params.LST[[MODE]]$space.target*100`

We solved each conservation planning data set using the unreliable and reliable problem formulations (using default failure distance and level parameter) using no boundary length modifer, and using a boundary length modifier of `r max(benchmark.params.LST[[MODE]]$blm)`. To assess variation in execution times, we generated and solved `r benchmark.params.LST[[MODE]]$replicates` random data sets per combination of number of features and planning units. All conservation planning problems were solved to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality using Gurobi (version 7.0.0) on a system with 64 GB RAM and an Intel(R) Xeon(R) CPU (E5-2630 v3; 2.40 GHz). All problems were solved using a single thread.

@@

The benchmark analysis revealed that moderately-sized conservation problems can be solved in a realtively short period of time (Figure 1). As predicted, problems were solved much quicker when using the unreliable formulation than when using the reliable formulation. Additionally, the solve times varied substiantially for different data sets with the same proporties (ie. number of features and planning units). This result suggests that it may difficult to estimate exactly how long a given problem will take to solve.

```{r, echo=FALSE, message=FALSE, fig.height=5.0, fig.width=4.5, fig.cap='Benchmark analysis. Boxplots show the amount of time required to solve a problem under a given set of parameters. Note that times are shown on a log_{10} scale.'}
# prepare data
plot.DF <- benchmark.DF %>%
  mutate(blm = paste0('BLM = ',blm)) %>% 
  mutate(number.features=factor(number.features)) %>%
  mutate(formulation = paste0(formulation, ' formulation')) %>%
  mutate(number.planning.units=factor(number.planning.units))
# make plot
ggplot(plot.DF, aes(x=number.planning.units, y=time, color=number.features)) +
  geom_boxplot() +
  facet_grid(blm ~ formulation ) +
  scale_y_continuous(trans='log10') +
  xlab('Number of planning units') + ylab('Solve time (s)') +
  labs(color='Number of features') +
  theme(axis.line.x=element_line(), axis.line.y=element_line(),
    strip.background = element_rect(fill='grey20'),
    strip.text = element_text(color='white', size=12),
    legend.position='bottom')
```

We were not able to explore how long large-sized problems would take to solve because our computer system did not have enough memory. One method to reduce the problem size would be to simply use less demand points. Another method to reduce problem size would be to set a maximum distance between demand points and planning units such that demand points cannot be represented by planning units that that are too far away. Both of these method could be applied using functions provided in the \texttt{raptr R} package. Future research is needed to explore these methods and understand how the impact solution quality. Additionally, future research could involve implementing more advanced methods to solve the optimization problem [eg. as described in @r16; @r15].

# Figures 
```{r, echo=FALSE, fig.height=8, fig.width=8.5, fig.cap='Attribute spaces used in the first case-study. Each panel shows a the distribution of a solution in environmental space and how it samples the realized niche for a different species. The left column of panels shows the solution generated using amount targets. The right column of panels shows the solution generated using amount and niche targets. Each column of panels corresponds to a different species. Hexagons show the distribution of demand points. The color of each heaxgon denotes the weighted frequency of demand points inside it. Points represent the environmental conditions associated with planning units inside the species geographic range that were selected for preservation in a given solution.'}
## prepare data
cs1.pu.niche.DF <- ddply(
	cs1.spp.DF,
	c('Species', 'Prioritisation'),
	function(x) {
		# init
		curr.prioritisation <- cs1.prioritisations[[as.character(x$Prioritisation)]]
		curr.spp.pos <- match(x$Species[1], curr.prioritisation@data@species[[1]])
		curr.sel.pos <- which(curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@ids %in% which(as.logical(selections(curr.prioritisation))))
		# extract planning units
		curr.DF <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@coords %>% as.data.frame()
		names(curr.DF) <- c('Niche1', 'Niche2')
		curr.DF$Species <-x$Species[[1]]
		curr.DF$Prioritisation <-x$Prioritisation[[1]]
		curr.DF$Solution <- 'discarded'
		curr.DF$Solution[curr.sel.pos] <- 'selected'
		# return data
		return(curr.DF)
	}
)
cs1.pu.niche.DF <- filter(cs1.pu.niche.DF, Solution=='selected')

cs1.dp.niche.DF <- ddply(
	cs1.spp.DF,
	c('Species', 'Prioritisation'),
	function(x) {
		# init
		curr.prioritisation <- cs1.prioritisations[[as.character(x$Prioritisation)]]
		curr.spp.pos <- match(x$Species[1], curr.prioritisation@data@species[[1]])
		# extract demand points
		curr.DF <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@demand.points@coords %>% as.data.frame()
		names(curr.DF) <- c('Niche1', 'Niche2')
		curr.DF$weight <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@demand.points@weights
		# return data
		return(curr.DF)
	}
)
cs1.pu.niche.DF$Prioritisation <- factor(as.character(cs1.pu.niche.DF$Prioritisation), levels=names(cs1.prioritisations))
cs1.dp.niche.DF$Prioritisation <- factor(as.character(cs1.dp.niche.DF$Prioritisation), levels=names(cs1.prioritisations))

# make letter data.frame
cs1.letters.DF <- expand.grid(
	Species=levels(cs1.pu.niche.DF$Species),
	Prioritisation=levels(cs1.pu.niche.DF$Prioritisation)
) %>% arrange(Species, Prioritisation)
cs1.letters.DF$letter <- paste0('(', letters[seq_len(nrow(cs1.letters.DF))], ')')
cs1.letters.DF <- cbind(
	cs1.letters.DF,
	rbind(
		cs1.pu.niche.DF %>% select(Species, Prioritisation, Niche1, Niche2),
		cs1.dp.niche.DF %>% select(Species, Prioritisation, Niche1, Niche2)
	) %>% summarise(
		Niche1=min(Niche1)+(abs(diff(range(Niche1)))*0.01),
		Niche2=min(Niche2)+(abs(diff(range(Niche2)))*0.95)
	) %>% data.frame()
)

## make plots
ggplot() +
	stat_summary_hex(data=cs1.dp.niche.DF, aes(x=Niche1, y=Niche2, z=weight), bins=15, fun=function(z) {sum(z)}) +
	geom_point(data=cs1.pu.niche.DF, aes(x=Niche1, y=Niche2), fill='#00441b', pch=21, color='black') +
	geom_text(data=cs1.letters.DF, aes(x=Niche1, y=Niche2, label=letter), color='black') +
	theme_classic() +
	theme(
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white')
	) +
	xlab('Niche (axis 1)') + ylab('Niche (axis 2)') +
	scale_fill_gradientn(name='Total demand\npoint weight', colors=colorRampPalette(brewer.pal(6, 'PuBu')[-1])(100)) +
	facet_grid(Species ~ Prioritisation)
```

@@

```{r, echo=FALSE, results='asis'}
# make table
term.CHR <- c(
	'$F$', 'set of biodiversity features (indexed by $f$)',
	'$J$', 'set of planning units (indexed by $j$)',
	'$q_{fj}$', 'probability of feature $f$ occupying planning unit $j$',
	'$e_{jk}$', 'shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrize exposed edges with no neighbours.',
	'$S$', 'set of attribute spaces (indexed by $S$)',
	'$I_{fsi}$', 'set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$',
	'$\\lambda_{fsi}$', 'set of weights for demand point $i$ for feature $f$ in attribute space $s$',
	'$d_{fsij}$', 'distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$',
	'$\\delta_{fsi}$', 'the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$',
	'$T_f$', 'amount target for feature $f$',
	'$\\tau_{fs}$', 'space-based target for feature $f$ in attribute space $s$',
	'$X_j$', 'binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0)',
	'$Y_{fsij}$', 'binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units'
)
term.DF <- data.frame(
	Symbol=term.CHR[seq(1, length(term.CHR), 2)],
	Description=term.CHR[seq(2, length(term.CHR), 2)]
)
term.DF$Description <- paste0('\\parbox[t][][t]{15cm}{', term.DF$Description, '}')

# render table
curr.file <- tempfile(fileext='.tex')
tmp <- latex(
	term.DF,
	file=curr.file, 
	digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
	colheads=c('Symbol', 'Description'), col.just=c('c', 'l'),
	caption='Symbols and descriptions of terms used in the formulation of the unreliable representative and adequate prioritization (URAP) problem.'
)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```

>>>>

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir = normalizePath("../.."))
```

@@

```{r, include = FALSE}
## initialization
# load session
try(session::restore.session("data/final/results.rda"))

## preliminary processing
# redefine file paths for rasters
for (i in seq_len(raster::nlayers(cs1_bioclim_raster)))
  cs1_bioclim_raster@layers[[i]]@file@name  <- paste0(getwd(),
    strsplit(cs1_bioclim_raster@layers[[i]]@file@name,
             "raptr-manuscript")[[1]][[2]])

@@

The reliable formulation (RRAP) is a multi-objective optimization problem.

@@

We generated simulated conservation planning data sets using varying number of features (`r paste(benchmark_parameters$number_of_features, collapse = ",")`) and planning units (`r paste(benchmark_parameters$number_of_planning_units, collapse = ",")`). For a given number of planning units and features, the data was simulated using the following steps. First, the planning units were simulated as squares that were arranged in a grid. All planning unit costs were set to one. The spatial distribution of features were generated using binary, spatially autocorrelated random fields [using the \texttt{RandomFields R} package; @r494; @r493]. If features were associated with less than two occupied planning units, these features were set to occupy two randomly selected planning units. All occupied planning units were set to have a probability of occupancy of `r benchmark_parameters$occupancy_probability`. Similar to the simulated data sets used in the main text, a geographic attribute space was used for each feature. The features' demand points were set as the centroids of the planning units that they occupied, and were weighted by the units' probability of occupancy. All problems used amount-based targets of `r benchmark_parameters$amount_target*100` % and space-based targets of `r benchmark_parameters$space_target*100`

We solved each conservation planning data set using the unreliable and reliable problem formulations (using default failure distance and level parameter) using no boundary length modifier, and using a boundary length modifier of `r max(benchmark_parameters$blm)`. To assess variation in execution times, we generated and solved `r benchmark_parameters$replicates` random data sets per combination of number of features and planning units. All conservation planning problems were solved to within `r general_parameters$MIPGap * 100` % of optimality using Gurobi (version 7.0.0) on a system with 64 GB RAM and an Intel(R) Xeon (R) CPU (E5-2630 v3; 2.40 GHz). All problems were solved using a single thread.

@@

The benchmark analysis revealed that moderately-sized conservation problems can be solved in a relatively short period of time (Figure S2). As predicted, problems were solved much quicker when using the unreliable formulation than when using the reliable formulation. Additionally, the solve times varied substantially for different data sets with the same properties (ie. number of features and planning units). This result suggests that it may difficult to estimate exactly how long a given problem will take to solve.

We were not able to explore how long large-sized problems would take to solve because our computer system did not have enough memory. One method to reduce the problem size would be to simply use less demand points. Another method to reduce problem size would be to set a maximum distance between demand points and planning units such that demand points cannot be represented by planning units that that are too far away. Both of these method could be applied using functions provided in the \texttt{raptr R} package. Future research is needed to explore these methods and understand how the impact solution quality. Additionally, future research could involve implementing more advanced methods to solve the optimization problems [eg. as described in @r16; @r15].

\clearpage

# Figures
```{r, echo = FALSE, fig.height = 8, fig.width = 8.5, fig.cap = "Attribute spaces used in the first case-study. Each panel shows a the distribution of a solution in environmental space and how it samples the realized niche for a different species. The left column of panels shows the solution generated using amount targets. The right column of panels shows the solution generated using amount and niche targets. Each column of panels corresponds to a different species. Hexagons show the distribution of demand points. The color of each hexagon denotes the weighted frequency of demand points inside it. Points represent the environmental conditions associated with planning units inside the species geographic range that were selected for preservation in a given solution."}
## prepare data
cs1_pu_niche_data <- plyr::ddply(cs1_spp_results,
  c("Species", "Prioritisation"),
  function(x) {
    # init
    curr_prioritisation <- cs1_prioritisations[[as.character(x$Prioritisation)]]
    curr_spp_pos <- match(x$Species[1], curr_prioritisation@data@species[[1]])
    curr_ids <- curr_prioritisation %>%
                slot("data") %>%
                slot("attribute.spaces") %>%
                `[[`(1) %>%
                slot("spaces") %>%
                `[[`(curr_spp_pos) %>%
                slot("planning.unit.points") %>%
                slot("ids")
    curr_sel_pos <- which(curr_ids %in% which(as.logical(
                          raptr::selections(curr_prioritisation))))
    # extract planning units
    curr_data <- curr_prioritisation %>%
                slot("data") %>%
                slot("attribute.spaces") %>%
                `[[`(1) %>%
                slot("spaces") %>%
                `[[`(curr_spp_pos) %>%
                slot("planning.unit.points") %>%
                slot("coords") %>%
                as.data.frame() %>%
                magrittr::set_names(c("Niche1", "Niche2")) %>%
                dplyr::mutate(Species = x[["Species"]][[1]],
                              Prioritisation = x[["Prioritisation"]][[1]],
                              Solution <- "discarded")
    curr_data$Solution[curr_sel_pos] <- "selected"
    # return data
    return(curr_data)
})
cs1_pu_niche_data %<>% dplyr::filter(Solution == "selected")

cs1_dp_niche_data <- plyr::ddply(
  cs1_spp_results,
  c("Species", "Prioritisation"),
  function(x) {
    # init
    curr_prioritisation <- cs1_prioritisations[[as.character(x$Prioritisation)]]
    curr_spp_pos <- match(x$Species[1], curr_prioritisation@data@species[[1]])
    # extract demand points
    curr_data <- curr_prioritisation %>%
                  slot("data") %>%
                  slot("attribute.spaces") %>%
                  `[[`(1) %>%
                  slot("spaces") %>%
                  `[[`(curr_spp_pos) %>%
                  slot("demand.points") %>%
                  slot("coords") %>%
                  as.data.frame() %>%
                  magrittr::set_names(c("Niche1", "Niche2"))
    curr_data$weight <- curr_prioritisation %>%
                        slot("data") %>%
                        slot("attribute.spaces") %>%
                        `[[`(1) %>%
                        slot("spaces") %>%
                        `[[`(curr_spp_pos) %>%
                        slot("demand.points") %>%
                        slot("weights")
    # return data
    return(curr_data)
})
cs1_pu_niche_data$Prioritisation <- factor(as.character(
                                      cs1_pu_niche_data$Prioritisation),
                                      levels = names(cs1_prioritisations))
cs1_dp_niche_data$Prioritisation <- factor(as.character(
                                      cs1_dp_niche_data$Prioritisation),
                                      levels = names(cs1_prioritisations))

# make letter data.frame
cs1_letters_data <- expand.grid(Species = levels(cs1_pu_niche_data$Species),
                                Prioritisation = levels(
                                  cs1_pu_niche_data$Prioritisation)) %>%
                    dplyr::arrange(Species, Prioritisation) %>%
                    dplyr::mutate(letter = paste0("(",
                      letters[seq_len(nrow(.))], ")"))

cs1_letters_data2 <- cs1_pu_niche_data %>%
                     dplyr::select(Species, Prioritisation, Niche1, Niche2) %>%
                     rbind(cs1_dp_niche_data %>% dplyr::select(Species,
                                                               Prioritisation,
                                                               Niche1,
                                                               Niche2)) %>%
                     dplyr::summarise(Niche1 = min(Niche1) +
                                               (abs(diff(range(Niche1))) *
                                               0.01),
                                      Niche2 = min(Niche2) +
                                               (abs(diff(range(Niche2))) *
                                               0.95)) %>%
                     data.frame()

cs1_letters_data %<>% cbind(cs1_letters_data2)

## make plots
ggplot2::ggplot() +
  ggplot2::stat_summary_hex(data = cs1_dp_niche_data,
                            ggplot2::aes(x = Niche1, y = Niche2, z = weight),
                            bins = 15, fun = function(z) sum(z)) +
  ggplot2::geom_point(data = cs1_pu_niche_data,
                      ggplot2::aes(x = Niche1, y = Niche2),
                      fill = "#00441b", pch = 21, color = "black") +
  ggplot2::geom_text(data = cs1_letters_data,
                     ggplot2::aes(x = Niche1, y = Niche2, label = letter),
                     color = "black") +
  ggplot2::theme_classic() +
  ggplot2::theme(strip.background = ggplot2::element_rect(fill = "grey20"),
                 strip.text = ggplot2::element_text(color = "white")) +
  ggplot2::xlab("Niche (axis 1)") +
  ggplot2::ylab("Niche (axis 2)") +
  ggplot2::scale_fill_gradientn(name = "Total demand\npoint weight",
                                colors = colorRampPalette(
                                  RColorBrewer::brewer.pal(6,
                                                          "PuBu")[-1])(100)) +
  ggplot2::facet_grid(Species ~ Prioritisation)
```

```{r, echo = FALSE, message = FALSE, fig.height = 5.0, fig.width = 4.5, fig.cap = "Time required to solve different sized problems using different parameters and formulations. Points represent average times and bars show standard errors. Time is shown on a log$_{10}$ scale."}
# prepare data
plot_data <- benchmark_results %>%
  dplyr::mutate(blm = paste0("BLM = ", blm)) %>%
  dplyr::mutate(number_features = factor(number_features)) %>%
  dplyr::mutate(formulation = paste0(formulation, " formulation")) %>%
  dplyr::mutate(number_planning_units = factor(number_planning_units))
# make plot
ggplot2::ggplot(plot_data,
                ggplot2::aes(x = number_planning_units, y = time,
                             color = number_features)) +
  ggplot2::stat_summary(fun.y = mean, geom = "point",
                        position = ggplot2::position_dodge(0.95)) +
  ggplot2::stat_summary(fun.data = ggplot2::mean_se, geom = "errorbar",
                        position = ggplot2::position_dodge(0.95)) +
  ggplot2::facet_grid(blm ~ formulation ) +
  ggplot2::scale_y_continuous(trans = "log10") +
  ggplot2::xlab("Number of planning units") +
  ggplot2::ylab("Solve time (s)") +
  ggplot2::labs(color = "Number of features") +
  ggplot2::theme(
    axis.line.x = ggplot2::element_line(),
    axis.line.y = ggplot2::element_line(),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    legend.position = "bottom")
```

\clearpage

@@

```{r, echo = FALSE, results = "asis"}
# make table
terms <- c(
  "$F$", "set of biodiversity features (indexed by $f$)",
  "$J$", "set of planning units (indexed by $j$)",
  "$q_{fj}$", "probability of feature $f$ occupying planning unit $j$",
  "$e_{jk}$", "shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrize exposed edges with no neighbours.",
  "$S$", "set of attribute spaces (indexed by $S$)",
  "$I_{fsi}$", "set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$",
  "$\\lambda_{fsi}$", "set of weights for demand point $i$ for feature $f$ in attribute space $s$",
  "$d_{fsij}$", "distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$",
  "$\\delta_{fsi}$", "the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$",
  "$T_f$", "amount target for feature $f$",
  "$\\tau_{fs}$", "space-based target for feature $f$ in attribute space $s$",
  "$X_j$", "binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0)",
  "$Y_{fsij}$", "binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determining the amount of the attribute space sampled by the selected planning units"
)
term_data <- data.frame(
  Symbol = terms[seq(1, length(terms), 2)],
  Description=terms[seq(2, length(terms), 2)]
)
term_data$Description <- paste0("\\parbox[t][][t]{15cm}{",
                                term_data$Description, "}")

# render table
curr_file <- tempfile(fileext = ".tex")
tmp <- Hmisc::latex(term_data,
                    file = curr_file,
                    digits = 2, rowname = NULL, first.hline.double = FALSE,
                    booktabs = TRUE, here = TRUE,
                    colheads = c("Symbol", "Description"),
                    col.just = c("c", "l"),
                    caption = paste0("Symbols and descriptions of terms used ",
                                     "in the formulation of the unreliable ",
                                     "representative and adequate ",
                                     "prioritization (URAP) problem."))
curr_file %>%
  readLines() %>%
  `[`(-1) %>%
  paste0(collapse = "\n") %>%
  gsub(pattern = "toprule", replacement = "toprule[1pt]", fixed = TRUE) %>%
  gsub(pattern = "bottomrule", replacement = "bottomrule[1pt]",
       fixed = TRUE) %>%
  cat()
```

\clearpage
